% =============================================================================
%  DRAFT-ONLY CHAPTER -- DELETE BEFORE FINAL SUBMISSION
%
%  This chapter is a working punch-list of items still outstanding on the
%  dissertation. It is included in main.tex via a single \include line and
%  is meant to be removed when the document is submitted. To remove it:
%
%    1.  Comment out (or delete) the line `\include{chapters/todo_tracker}`
%        in main.tex.
%    2.  Optionally delete this file entirely.
%
%  Do not cite anything from this chapter elsewhere in the dissertation:
%  no \label here is referenced from any other chapter, by design.
% =============================================================================
\chapter*{TODO --- Outstanding Work (Draft Only)}
\thispagestyle{anonymous}
\addcontentsline{toc}{chapter}{TODO --- Outstanding Work (Draft Only)}

\begin{center}
    \fbox{\parbox{0.9\linewidth}{%
        \centering\bfseries
        This chapter must be removed before final submission. See the
        comment block at the top of \texttt{chapters/todo\_tracker.tex}
        for the one-line change in \texttt{main.tex} that disables it.
    }}
\end{center}

\section*{Personal Front-Matter Fields}

\begin{itemize}
    \item Set \texttt{\textbackslash studentname} in \texttt{main.tex}
        (currently empty so the marker-blind pages compile cleanly).
    \item Set \texttt{\textbackslash studentid} in \texttt{main.tex}
        (currently the placeholder \texttt{201600000}).
    \item Set \texttt{\textbackslash supervisorname} in
        \texttt{main.tex} (currently \texttt{TODO Supervisor Name}).
    \item Set \texttt{\textbackslash moduleassessor} in
        \texttt{main.tex} (currently
        \texttt{TODO Second Marker / Assessor}).
    \item Verify the data and participant category codes
        (\texttt{A0/A0}) in \texttt{chapters/ethical\_compliance.tex}
        against the latest COMP390 handbook.
\end{itemize}

\section*{Figures Still on Placeholders}

\begin{itemize}
    \item \texttt{figures/architecture.pdf} for
        \Cref{fig:arch} in \Cref{ch:design} --- currently an ASCII
        \texttt{\textbackslash fbox} placeholder. Draw the
        ViZDoom~$\rightarrow$~COOM~$\rightarrow$~SAC pipeline in TikZ
        or export from draw.io; expected size $\sim 0.85$\,linewidth.
\end{itemize}

\section*{Second-Stage Algorithmic Work}

The supervisor's mid-project feedback set the second-stage scope
(\Cref{sec:scope_update}). Items below execute that scope.

\begin{itemize}
    \item Implement \texttt{CL/methods/ewc\_online.py}: subclass
        \texttt{Regularization\_SAC}, override \texttt{\_merge\_weights}
        with the EMA rule
        $F \leftarrow \gamma F_{\text{old}} + (1 - \gamma) F_{\text{new}}$.
        Add \texttt{--ewc\_gamma} to \texttt{config.py}, default 0.95.
        Register \texttt{ewc\_online} in the \texttt{CLMethod} enum
        in \texttt{run\_cl.py}.
    \item Optionally also implement
        \texttt{CL/methods/ewc\_layerwise.py} (per-layer~$\lambda$ for
        CNN encoder, shared MLP, task-specific heads).
    \item Smoke-test the new method on CO4 / $20{,}000$ steps /
        seed~1 and verify \texttt{progress.tsv} contains a non-zero
        \texttt{train/loss\_reg} from task~1 onward.
\end{itemize}

\section*{Second-Stage Experimental Runs}

\begin{itemize}
    \item \emph{Improved EWC vs original EWC head-to-head.} Run
        \texttt{ewc\_online} (and/or \texttt{ewc\_layerwise}) at
        CO4 / $20{,}000$ for seeds $\{1,2,3\}$. Original EWC at the
        same seeds is already on disk and can be reused.
    \item \emph{Cross-family baselines.} Run ClonEx-SAC at CO4 /
        $20{,}000$ for seeds $\{1,2,3\}$. Run PackNet at CO4 /
        $20{,}000$ for seeds $\{1,2,3\}$.
    \item \emph{Backfill seed band on R3.} Run Fine-Tuning, L2 and
        MAS at seeds 2 and 3 to close requirement~\ref{req:variance}
        fully.
    \item \emph{Held-out evaluation pass.} Re-run at least one EWC
        and one improved-EWC seed with \texttt{--test} enabled, to
        produce true Forgetting and Forward Transfer numbers (the
        present study reports training-time success only;
        \Cref{sec:validity}).
\end{itemize}

\section*{Optional Cloud Replication}

\begin{itemize}
    \item Rent a single RTX 3090 on RunPod ($\sim$\$15--30) and
        replicate the EWC vs improved-EWC head-to-head at the
        published budget of $200{,}000$ steps per task with three
        seeds. This is the most direct test of the
        Fisher-on-noise diagnosis (\Cref{sec:fishernoise}) and the
        item~4 in the future-work list (\Cref{sec:future}).
\end{itemize}

\section*{Writing Backfill After Second-Stage Runs}

\begin{itemize}
    \item Add an algorithm description of the improved EWC
        variant(s) to \Cref{ch:impl}, with a small code listing.
    \item Add a results subsection to \Cref{ch:eval} for the
        head-to-head against original EWC (table + curve figure).
    \item Add a results subsection to \Cref{ch:eval} for the
        cross-family baseline comparison (extend
        \Cref{tab:avg_perf} with ClonEx and PackNet rows).
    \item Adjust the abstract and the \Cref{ch:intro} narrative
        from a pure first-stage negative-result framing to a
        ``negative result $\rightarrow$ targeted fix $\rightarrow$
        improved variant'' arc, once the second-stage numbers are
        in.
    \item Update the requirements mapping in
        \Cref{sec:reqs_met}: once Fine-Tuning, L2 and MAS have
        three-seed runs, R3 changes from ``partially met'' to
        ``met''.
\end{itemize}

\section*{House-Keeping Before Submission}

\begin{itemize}
    \item Re-run \texttt{paper/scripts/make\_figures.py} after
        every new run is added, and re-copy the outputs into
        \texttt{dissertation/figures/} and
        \texttt{dissertation/tab*.tex}.
    \item Squash the five overfull-\texttt{hbox} warnings reported
        by \texttt{tectonic} (long URLs and compound words in
        \Cref{ch:intro}, \Cref{ch:design},
        \Cref{ch:eval} and \Cref{ch:concl}).
    \item Verify every \texttt{\textbackslash citep} resolves
        without a \texttt{??}{} marker in the rendered PDF.
    \item \textbf{Delete this chapter:} comment out
        \texttt{\textbackslash include\{chapters/todo\_tracker\}}
        in \texttt{main.tex} and re-run \texttt{tectonic main.tex}.
\end{itemize}
